---
title: "LSKVC Return to Play (Indoor) Players Rules"
author: "Nicolas Vecchione (Powerball Rep)"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
  word_document:
    toc: true
    toc_depth: 2
    number_sections: true
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r source_functions, echo=FALSE,message=FALSE, warning=FALSE}
source("ProduceConfigurations.R")
```

# Motivation {-}

This document aims to support LSKVC in their organisation and running of official LSKVC Indoor volleyball sessions returning from the COVID19 hiatus. Its overarching aim is to provide means to reduce the risks of virus propagation, protecting our members and the general population and create the start of a sustainable way to play volleyball during the ongoing pandemic. 

The **Players Rules** section outlines behaviour expected from players during LSKVC Indoor training sessions. 

The  Player Rules are made of:

* Rules and obligations as extracted and adapted for our purposes from the Volleyball England document titled "Volleyball England Return to Indoor Volleyball Specific Action Plan" (as updated 13th May 2021 for step 3 of roadmap out of lockdown) ([link](https://media.volleyballengland.org/docs/Return%20to%20play%20indoor%20volleyball%20-%20sport%20specific%20action%20plan.pdf))
* Additional specific rules based on our specific Risk Assessment.

The **Risk Assessment** (available at  *[https://github.com/cuisquare/ReturnToVolleyball/blob/master/IndoorTraining/LSKVC_IndoorTraining_RiskAssessment.pdf](https://github.com/cuisquare/ReturnToVolleyball/blob/master/IndoorTraining/LSKVC_IndoorTraining_RiskAssessment.pdf)*) must be read and understood alongside those rules. 

The **Club Requirements** (available at  *[https://github.com/cuisquare/ReturnToVolleyball/blob/master/IndoorTraining/LSKVC_IndoorTraining_ClubRequirements.pdf](https://github.com/cuisquare/ReturnToVolleyball/blob/master/IndoorTraining/LSKVC_IndoorTraining_ClubRequirements.pdf)*) should also be read to be aware of the level of support you can expect during a session.

\newpage
# Players Rules

## Prior to Activity:
* When requested, let us know you plan to attend the activity, your contact details and confirm you opt-in to the present rules and risk assessment and agree with the residual risk involved in the activity. Note that if you have not provided the above information, you will not be able to attend. 
* Note that by attending you agree for your personal contact details to be stored and held for 21 days following the activity and will be supplied to NHS Track and Trace/ other relevant public authority if requested to do so.
* Numbers are capped at *17 players* and numbers will be managed by the session organiser. You will receive confirmation of your spot most likely through a Teamer invite but your individual team might do things differently. Do not come over without that confirmation.
* Remember to follow all current UK Government guidances (see https://www.gov.uk/coronavirus). This includes most importantly not taking part if you experience any symptoms of Covid19 as well as maintaining social distancing in making your way to the venue.
* Bring your own food and drink – water bottles or other refreshment containers should not be shared. 
* Arrive to the venue already changed and carry out finishing touches of changing on court.
* Make yourself aware of club specific risk assessment (https://github.com/cuisquare/ReturnToVolleyball/blob/master/IndoorTraining/LSKVC_IndoorTraining_RiskAssessment.pdf) that those rules are based on
* Specific indoor venues will have their own risk assessment and rules, which you will have to make yourself aware of when communicated to you. This might include access to the venue, use of changing facilities and access times.

## During Activity:
* Follow the additional rules provided by your specific venue.
* Arrive to the venue already changed and carry out finishing touches of changing on court. Only use changing facilities to access toilets/hand washing. 
* Sanitise hands before handling training equipment for first time.
* Sanitise hands immediately prior to entering the court and immediately after exiting the court, including for water breaks, or as directed by coach/covid officer (to avoid transfer of virus material between the training equipment and your personal belongings).
* Congratulatory touches are not permitted (if you do one by mistake, please acknowledge and do comply with covid officer reminder). 
* Players should not to touch their face, mouth, nose or eyes. 
* Shouting should be avoided (shouting produces aerosol droplets which propagates virus longer and further than normal breathing)
* Any participants not on court must maintain social distancing. Any seating arrangement (where you keep your stuff) should be 2m away from the next persons. 
* Water and drinks should not be shared by people from different households

## After the Activity:
* Sanitise your hands after the completion of activity
* Leave the playing area whilst maintaining social distancing.
* If you develop symptoms and are asked to supply contacts to NHS Track and Trace, the person to contact is the covid officer on the day (See your confirmation email for their details). If you can't place them, please contact your team rep
* If you have suggestions on running the session or observed something you think could be improved especially in terms of safety please get in touch with the organisers
